Report for Qualifying Exam

### 2024-08-15 src/main.tex
completed related work section w/o table (due review)
changes in related work, table pending
added related work section, pending:table citation, caption
modified the table, added citations
worked on RP1 section
added context for hamilton paper
completed section 2 (DONE :) )
minor changes
changed title, modified RP1
report writing done, conclusion pending
added conclusion section, minor changes
minor edits before review

\documentclass[manuscript, screen, acmsmall]{acmart}
\usepackage[noend]{algorithm2e}
\usepackage{graphicx,balance,multirow,xspace,subcaption,longtable}
\usepackage{tabularx,multirow,booktabs,blindtext}
\usepackage{makecell,amsmath,cleveref,rotating}
\usepackage[font={small}, textfont=md]{caption}
\usepackage[utf8]{inputenc}
\usepackage[draft,inline,nomargin,index]{fixme}
\usepackage{marginnote}
\usepackage[bb=boondox,bbscaled=.95,cal=boondoxo]{mathalfa}
\usepackage{adjustbox}
\usepackage{colortbl}

@@

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref} \urlstyle{tt}
\newcolumntype{P}[1]{>{\arraybackslash}p{#1}}
%\newcolumntype{X}[1]{>{\centering\arraybackslash}p{#1}}

\crefformat{section}{\S#2#1#3}
\crefformat{subsection}{\S#2#1#3}
\crefformat{subsubsection}{\S#2#1#3}

\hypersetup{%
  pdftitle={Report for Qualifying Exam},
  pdfauthor={},
  pdfkeywords={},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

@@

\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}

@@

% \renewcommand{\footnotesize}{\scriptsize}
\newcommand{\etc}{etc.}
\newcommand{\eg}{e.g.,\ }
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.,\ }
\newcommand{\re}{r.e.\ }
\newcommand{\aka}{a.k.a.\ }
\newcommand{\cf}{\emph{cf.\ }}
\newcommand{\note}[1]{\textcolor{blue}{#1}}
\newcommand{\superem}[1]{\underline{\emph{#1}}}
\newcommand{\para}[1]{\vspace{.05in}\noindent\textbf{#1}}
\newcommand{\paraem}[1]{\vspace{.05in}\noindent\emph{#1}}
\newcommand{\yt}{YouTube\xspace}
\newcommand{\X}{$\mathbb{X}$\xspace}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}

\begin{document}

\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers

\renewcommand{\sectionautorefname}{\S}
\renewcommand{\subsectionautorefname}{\S}
\renewcommand{\subsubsectionautorefname}{\S}

\input{sections/titlepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%% Table of Contents
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document
%\maketitle
%\sloppy
\input{sections/abstract}
\input{sections/introduction}
\newpage
\input{sections/relatedwork}
\newpage
\input{sections/papers}
\balance

\newpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

\end{document}

>>>>

\documentclass[screen, acmsmall]{acmart}
\usepackage[noend]{algorithm2e}
\usepackage{graphicx,balance,multirow,xspace,subcaption,longtable}
\usepackage{tabularx,multirow,booktabs,blindtext}
\usepackage{makecell,amsmath,cleveref,rotating}
\usepackage[font={small}, textfont=md]{caption}
\usepackage[utf8]{inputenc}
\usepackage[draft,inline,nomargin,index]{fixme}
\usepackage{marginnote}
\usepackage[bb=boondox,bbscaled=.95,cal=boondoxo]{mathalfa}
\usepackage{adjustbox}
\usepackage{colortbl}
\usepackage[table,xcdraw]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{wrapfig}
\usepackage{svg}
\useunder{\uline}{\ul}{}

@@

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{wasysym} \urlstyle{tt}
%\newcolumntype{P}[1]{>{\arraybackslash}p{#1}}
%\newcolumntype{X}[1]{>{\centering\arraybackslash}p{#1}}

%\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
%  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
%  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
%  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
%  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
%  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
%  \do\Y\do\Z}

%\crefformat{section}{\S#2#1#3}
%\crefformat{subsection}{\S#2#1#3}
%\crefformat{subsubsection}{\S#2#1#3}

\hypersetup{%
  pdftitle={Report for Qualifying Exam},
  pdfauthor={Manisha Keim},
  pdfkeywords={},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

@@

%\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
%\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}

@@

% \renewcommand{\footnotesize}{\scriptsize}
\newcommand{\etc}{etc.}
\newcommand{\eg}{e.g.,\ }
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.,\ }
\newcommand{\re}{r.e.\ }
\newcommand{\aka}{a.k.a.\ }
\newcommand{\cf}{\emph{cf.\ }}
\newcommand{\note}[1]{\textcolor{blue}{#1}}
\newcommand{\superem}[1]{\underline{\emph{#1}}}
\newcommand{\para}[1]{\vspace{.05in}\noindent\textbf{#1}}
\newcommand{\paraem}[1]{\vspace{.05in}\noindent\emph{#1}}
\newcommand{\yt}{YouTube\xspace}
\newcommand{\X}{$\mathbb{X}$\xspace}
%\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}

\begin{document}
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers
\renewcommand{\sectionautorefname}{\S}
\renewcommand{\subsectionautorefname}{\S}
\renewcommand{\subsubsectionautorefname}{\S}

\input{sections/titlepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%% Table of Contents
\tableofcontents
%\input{sections/abstract}
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document
%  \title{Understanding the flow of Temporal Context Shift in Data Voids}
%  \maketitle
%  \sloppy

%\newpage
\input{sections/introduction}
\newpage
\input{sections/relatedwork}
\newpage
\input{sections/paper_kulkarni}
\newpage
\input{sections/paper_hamilton}
\newpage
\input{sections/conclusion}
\balance
\newpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}
\end{document}

### 2024-08-15 src/sections/relatedwork.tex
minor edits before review

\para{Semantic shifts with contextual embeddings:}
Giulianelli et al.\ (2020) explored the phenomenon of lexical semantic change using an unsupervised approach using \emph{BERT}.
They utilized the \emph{BERT} model to generate contextual word embeddings, which captured the meaning of a word based on its surrounding context.
The extracted word usage vectors are then clustered into different \emph{usage types} using the \emph{k-means clustering algorithm}.
This helped to identify distinct senses or meanings of the words as they appear in various contexts.
They analyzed these clusters for a specific word across different time periods.
The study proposed three metrics to quantify semantic change:
\begin{packed_enumerate}
    \item \textbf{Entropy Difference (ED)}: Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
    \item \textbf{Jensen-Shannon Divergence (JSD)}: Compares the similarity of word usage distributions across time intervals.
    \item \textbf{Average Pairwise Distance (APD)}: Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
\end{packed_enumerate}
The qualitative analysis indicated that the approach could capture various linguistic phenomena, including both synchronic (current usage) and diachronic (historical changes) aspects.

>>>>

\para{Semantic shifts with contextual embeddings:}
Giulianelli et al.\ (2020) explored the phenomenon of lexical semantic change using an unsupervised approach using \emph{BERT}.
They utilized the \emph{BERT} model to generate contextual word embeddings, which captured the meaning of a word based on its surrounding context.
The extracted word usage vectors are then clustered into different \emph{usage types} using the \emph{k-means clustering algorithm}.
This helped to identify distinct senses or meanings of the words as they appear in various contexts.
They analyzed these clusters for a specific word across different time periods.
The study proposed three metrics to quantify semantic change:
\begin{packed_enumerate}
    \item Entropy Difference (ED): Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
    \item Jensen-Shannon Divergence (JSD): Compares the similarity of word usage distributions across time intervals.
    \item Average Pairwise Distance (APD): Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
\end{packed_enumerate}
The qualitative analysis indicated that the approach could capture various linguistic phenomena, including both synchronic (current usage) and diachronic (historical changes) aspects.
